---
title: "Day 02B - PBMC Integration"
author: "Saket Choudhary <saketc@iitb.ac.in> for QSCB 2025"
date: today
format:
  html:
    embed-resources: true
    toc: false
    number-sections: true
    code-fold: show
    code-tools: true
    code-copy: true
    theme: cosmo
    css: styles.css
    fig-width: 8
    fig-height: 6
    df-print: paged
editor: visual
---

# Setup

```{r}
setRepositories(ind = 1:3)
required.packages <- c(
  "Seurat", "BiocManager",
  "harmony", "glmGamPoi", "devtools"
)
InstallMissing <- function(packages) {
  for (pkg in packages) {
    if (!require(pkg, character.only = TRUE)) {
      install.packages(pkg, dependencies = TRUE)
      library(pkg, character.only = TRUE)
    }
  }
}

BiocManager::install("batchelor")
library(Seurat)
library(dplyr)
devtools::install_github("satijalab/seurat-wrappers")
library(SeuratWrappers)
```

# Load data

```{r}
pbmc3k <- readRDS("/NAS/qscb2025/Guilliams_2022_livercellatlas/data/pbmc_integration/pbmc3k_annotated.rds")
pbmc3k$dataset <- "pbmc3k"
pbmc4k <- readRDS("/NAS/qscb2025/Guilliams_2022_livercellatlas/data/pbmc_integration/pbmc4k.rds")
pbmc4k$dataset <- "pbmc4k"

pbmc4k <- NormalizeData(pbmc4k) %>%
  FindVariableFeatures() %>%
  ScaleData() %>%
  RunPCA(npcs = 30) %>%
  FindNeighbors(dims = 1:20) %>%
  RunUMAP(dims = 1:20)
```

# Merge the two datasets

```{r}
merged <- merge(pbmc3k,
  y = pbmc4k,
  add.cell.ids = c("pbmc3k", "pbmc4k")
)
```

# Perform standard workflow for visualization and clustering

```{r}
merged <- JoinLayers(object = merged)

library(sparseMatrixStats)
df <- data.frame(
  mean = rowMeans(merged@assays$RNA@layers$counts),
  variance = rowVars(merged@assays$RNA@layers$counts)
)
merged <- NormalizeData(merged) %>%
  FindVariableFeatures(selection.method = "vst", nfeatures = 2000) %>%
  ScaleData() %>%
  RunPCA(npcs = 30) %>%
  FindNeighbors(dims = 1:20) %>%
  RunUMAP(dims = 1:20)
df <- data.frame(
  mean = rowMeans(merged@assays$RNA@layers$data),
  variance = rowVars(merged@assays$RNA@layers$data)
)
```

# Visualize the merged dataset

do the datasets merge into each other well?

```{r}
#| fig-width: 8
#| fig-height: 6

DimPlot(merged, reduction = "umap", group.by = "dataset")
```

```{r}
merged <- FindClusters(merged, resolution = 0.3)
merged$merged_cluster <- Idents(merged)
```

```{r}
#| fig-width: 8
#| fig-height: 6

DimPlot(merged, reduction = "umap", group.by = "merged_cluster")
```

# visualise the clusters

```{r}
#| fig-width: 8
#| fig-height: 6

DimPlot(merged, reduction = "umap", group.by = "merged_cluster", label = TRUE)
```

```{r}
#| fig-width: 10
#| fig-height: 5

DimPlot(merged, reduction = "umap", group.by = c(
  "dataset",
  "merged_cluster"
), label = TRUE)
```

# Let's integrate

```{r}
merged <- JoinLayers(merged)
merged[["RNA"]] <- split(merged[["RNA"]], f = merged$dataset)
merged <- NormalizeData(merged) %>%
  FindVariableFeatures() %>%
  ScaleData() %>%
  RunPCA()

integrated <- IntegrateLayers(
  object = merged,
  method = FastMNNIntegration,
  orig.reduction = "pca",
  new.reduction = "integrated.mnn",
  verbose = FALSE
)
integrated <- RunUMAP(integrated,
  reduction = "integrated.mnn", dims = 1:20,
  reduction.name = "umap.mnn"
)

integrated <- IntegrateLayers(
  object = integrated, method = HarmonyIntegration,
  orig.reduction = "pca", new.reduction = "integrated.harmony",
  verbose = FALSE
)


integrated <- IntegrateLayers(
  object = integrated, method = CCAIntegration,
  orig.reduction = "pca", new.reduction = "integrated.cca",
  verbose = FALSE
)

integrated <- FindNeighbors(integrated, reduction = "integrated.cca", dims = 1:20)
integrated <- FindClusters(integrated,
  resolution = 0.1,
  cluster.name = "cca_clusters"
)
integrated <- RunUMAP(integrated,
  reduction = "integrated.cca", dims = 1:20,
  reduction.name = "umap.cca"
)
```

```{r}
#| fig-width: 10
#| fig-height: 5

DimPlot(integrated,
  reduction = "umap.cca", group.by = c(
    "dataset",
    "merged_cluster", "cca_clusters"
  ),
  label = TRUE
)
```

```{r}
integrated <- FindNeighbors(integrated, reduction = "integrated.harmony", dims = 1:20)
integrated <- FindClusters(integrated,
  resolution = 0.3,
  cluster.name = "harmony_clusters"
)
integrated <- RunUMAP(integrated,
  reduction = "integrated.harmony", dims = 1:20,
  reduction.name = "umap.harmony"
)
```

```{r}
#| fig-width: 8
#| fig-height: 6

DimPlot(integrated)
```

```{r}
#| fig-width: 10
#| fig-height: 5

DimPlot(integrated,
  reduction = "umap.harmony", group.by = c(
    "dataset",
    "harmony_clusters", "cca_clusters"
  ),
  label = TRUE
)
```

# We can use the same idea to 'transfer' annotation from pbmc3k to pbmc4k

```{r}
pbmc3k <- NormalizeData(pbmc3k) %>%
  FindVariableFeatures(selection.method = "vst", nfeatures = 2000) %>%
  ScaleData() %>%
  RunPCA(npcs = 30) %>%
  FindNeighbors(dims = 1:20) %>%
  RunUMAP(dims = 1:20)
pbmc4k <- NormalizeData(pbmc4k) %>%
  FindVariableFeatures(selection.method = "vst", nfeatures = 2000) %>%
  ScaleData() %>%
  RunPCA(npcs = 30) %>%
  FindNeighbors(dims = 1:20) %>%
  RunUMAP(dims = 1:20)

anchors <- FindTransferAnchors(
  reference = pbmc3k,
  query = pbmc4k,
  dims = 1:30,
  reference.reduction = "pca"
)
predictions <- TransferData(
  anchorset = anchors, refdata = pbmc3k$seurat_annotations,
  dims = 1:30
)
pbmc4k <- AddMetaData(pbmc4k, metadata = predictions)
```

```{r}
#| fig-width: 8
#| fig-height: 6

DimPlot(pbmc4k, group.by = "predicted.id")
```

# uamp is only for visualisatio

always check using markers if the results make sense

```{r}
Idents(pbmc4k) <- pbmc4k$predicted.id
markers.B <- FindMarkers(object = pbmc4k, ident.1 = "B", only.pos = T, group.by = "predicted.id")
```
